Distinguishing among coalescent models using two-site allele frequency spectra

### 2018-08-07 manuscript/manuscript.tex

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/figure3.pdf}
\caption{Distortions in $\fpmi$ vs. distortions in the SFS for $n=4$. Parameters for beta coalescents and exponential growth are as in \fig{fig:fpmi_vs_tajimasd}. For the piecewise-constant--$N$ model, the fold-change in $N$ varies from 2 to $2^4$ and the time of this change varies from $2^-4$ to 2 in coalescent time units. \label{fig:sdpmi_vs_sdratio}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/figure5.pdf}
\caption{Pointwise mutual information between high-- and low--minor allele count mutations ($\hilopmi$) versus genetic distance. The first three rows show coalescent simulations in \texttt{msprime}. The fourth row shows forward-time \texttt{SLiM} simulations with selection at linked sites. Lines are colored by the distortion in the SFS, as measured by Tajima's D. [List parameters here?] \label{fig:hilo_vs_d}}
\end{figure}

@@

After filtering for coverage, removing chromosome arms with known inversions, downsampling to $n=100$ samples per autosomal chromosome arm, and identifying 4-fold degenerate sites, we selected the central region of each chromosome that have consistent high diversity (Methods).
Because the average pairwise diversity varies between arms---possibly reflecting selection or different sets of segregating inversions---we performed all subsequent calculations on each arm independently.
We fit a demographic model to the site frequency spectra of these central regions using \texttt{fastNeutrino} \cite{BhaskarEtal2015}.
We fit a 3-epoch piecewise-constant model, with four free parameters: two changepoints and two population size ratios.
We report our fitted parameters and those of a similar published model \autocite{RagsdaleGutenkunst2017} in Table~\ref{tab:DPGP_params} [Make table].
We then simulated under our fitted model using \texttt{msprime} and computed the expected and observed SFS, $\fpmi$, and $\hilopmi$ (\fig{fig:dpgp3}).

@@

In addition to the average SFS used to fit the model, we computed the average 2-SFS for pairs of sites at distances distances between 3~bp and 5~Kb.
Because we are using four-fold degenerate sites, we only computed the 2-SFS for distances that are multiples of 3.
A \texttt{snakemake} pipeline, \texttt{python} scripts, and \texttt{jupyter} notebooks to replicate our data processing, model fitting, simulations, and analysis of the DPGP3 data are available [GITHUB].

@@

\begin{table}[h!]
  \begin{center}
    \caption{[Table of parameters underlying Figure 2]}
    \label{tab:nonrec_params}
    \begin{tabular}{l|c|r} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
      % \textbf{Value 1} & \textbf{Value 2} & \textbf{Value 3}\\
      % $\alpha$ & $\beta$ & $\gamma$ \\
      % \hline
      % 1 & 1110.1 & a\\
      % 2 & 10.1 & b\\
      % 3 & 23.113231 & c\\
    \end{tabular}
  \end{center}
\end{table}

\begin{table}[h!]
  \begin{center}
    \caption{[Table of parameters underlying Figure 4]}
    \label{tab:n4_params}
    \begin{tabular}{l|c|r} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
      % \textbf{Value 1} & \textbf{Value 2} & \textbf{Value 3}\\
      % $\alpha$ & $\beta$ & $\gamma$ \\
      % \hline
      % 1 & 1110.1 & a\\
      % 2 & 10.1 & b\\
      % 3 & 23.113231 & c\\
    \end{tabular}
  \end{center}
\end{table}

\begin{table}[h!]
  \begin{center}
    \caption{[Table of parameters underlying Figure 6]}
    \label{tab:rec_params}
    \begin{tabular}{l|c|r} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
      % \textbf{Value 1} & \textbf{Value 2} & \textbf{Value 3}\\
      % $\alpha$ & $\beta$ & $\gamma$ \\
      % \hline
      % 1 & 1110.1 & a\\
      % 2 & 10.1 & b\\
      % 3 & 23.113231 & c\\
    \end{tabular}
  \end{center}
\end{table}

@@

TODO:
- Abstract
- Discussion
- Edit figures
- Parameter tables:
    - fig 2
    - fig 4
    - fig 6
- Get rest of citations
- Code:
    - organize
    - clean
    - comment
    - post on Github

>>>>

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/figure3.pdf}
\caption{Distortions in $\fpmi$ vs. distortions in the SFS for $n=4$. Parameters for beta coalescents and exponential growth are as in \fig{fig:fpmi_vs_tajimasd}. For the piecewise-constant--$N$ model, the fold-change in $N$ varies from 2 to $2^4$ and the time of this change varies from $2^{-4}$ to 2 in coalescent time units. \label{fig:sdpmi_vs_sdratio}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/figure5.pdf}
\caption{Pointwise mutual information between high-- and low--minor allele count mutations ($\hilopmi$) versus genetic distance. The first three rows show coalescent simulations in \texttt{msprime}. Exponential growth rate $g$ ranges from 0.01 to 8.0 in coalescent time units. Piecewise constant change-points, range from 0.01 to 1 in coalescent time units, and the ratio of ancestral to modern population size ranges from 0.01 to 0.2. The fourth row shows forward-time \texttt{SLiM} simulations with selective sweeps at linked sites (beneficial mutation fraction $f_{\text{sel}} \in \{10^{-6}, 10^{-5}, 10^{-4}, 10^{-3}\}$, see Methods). Lines are colored by the distortion in the SFS, as measured by Tajima's D. \label{fig:hilo_vs_d}}
\end{figure}

@@

After filtering for coverage, removing chromosome arms with known inversions, downsampling to $n=100$ samples per autosomal chromosome arm, and identifying 4-fold degenerate sites, we selected the central region of each chromosome that have consistent high diversity (Methods).
Because the average pairwise diversity varies between arms---possibly reflecting selection or different sets of segregating inversions---we performed all subsequent calculations on each arm independently.
We fit a demographic model to the site frequency spectra of these central regions using \texttt{fastNeutrino} \cite{BhaskarEtal2015}.
We fit a 3-epoch piecewise-constant model, with four free parameters: two changepoints and two population size ratios.
We report our fitted parameters in Table~\ref{tab:DPGP_params}.
We then simulated under our fitted model using \texttt{msprime} and computed the expected and observed SFS, $\fpmi$, and $\hilopmi$ (\fig{fig:dpgp3}, See methods for details.).

@@

In addition to the average SFS used to fit the model, we computed the average 2-SFS for pairs of sites at distances distances between 3~bp and 5~Kb.
Because we are using four-fold degenerate sites, we only computed the 2-SFS for distances that are multiples of 3.
For comparison between data and simulations, we scale the distances in basepairs by a critical distance $d_c = (r\E{T_2})^{-1}$.
We used a genome-wide recombination rate of $r = 2 \times 10^{-8}$ per basepair per generation \autocite{dmelrec}.
We estimated $\E{T_2}$ by $\pi / 2 \mu$.
For $\pi$, we used the average pairwise diversity at fourfold degenerate sites in the central region of each chromosome arm.
For $\mu$, we used a genome-wide mutation rate of $3 \times 10^{-9}$ per basepair per generation \autocite{dmelmu}.
These estimates are not precise, but only serve to scale genetic distances to the right order of magnitude.

A \texttt{snakemake} pipeline, \texttt{python} scripts, and \texttt{jupyter} notebooks to replicate our data processing, model fitting, simulations, and analysis of the DPGP3 data are available [GITHUB].

@@

@@

TODO:
- Abstract
- Discussion
- Edit figures
- Consistency of nomenclature, labeling, etc.
- Check drosophila numbers. Get citations or remake figure.
- Check references to Methods. (Should all be there and all have corresponding methods section.)
- Get rest of citations
- Code:
    - organize
    - clean
    - comment
    - post on Github
